Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics Experiments

### 2013-11-13 papers/2013/NSS/NSS.tex
make bibtex behave better
paragraph on OpenCL added

% disable the running header on all pages following the title page
\pagestyle{empty}
%
\begin{document}
% paper title
\title{Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics
Experiments}
% Author list
\author{A.~Gianelle, 
  S.~Amerio, 
  D.~Bastieri, 
  M.~Corvo, 
  W.~Ketchum,
  T.~Liu, 
  A.~Lonardo, 
  D.~Lucchesi,
  S.~Poprocki, 
  R.~Rivera, 
  L.~Tosoratto,
  P.~Vicini
  and 
  P.~Wittich
\thanks{Manuscript received November 15, 2013.
This work was supported by the  US National Science Foundation, the US 
Department of Energy Office of Science and the Italian
Istituto Nazionale di Fisica Nucleare. This work was partially supported by the 
EU Framework Programme 7 project EURETILE under grant number 247846.}% <-this % stops a space
\thanks{A.~Gianelle, S.~Amerio, D.~Bastieri, M.~Corvo, and D.~Lucchesi are with INFN Padova.}% 
\thanks{W.~Ketchum is with Los Alamos National Laboratory.}%
\thanks{T.~Liu and R.~Rivera are with Fermi National Accelerator Laboratory.}%
\thanks{A.~Lonardo, L.~Tosoratto and P.~Vicini are with INFN Roma}% 
\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}%
}
% make the title area
\maketitle
% after \maketitle, add the following command to remove the header on the title page
\thispagestyle{empty}
% abstract usually in passive voice
\begin{abstract}
  Interest in parallel architectures applied to real time selections is 
  growing in High Energy Physics (HEP) experiments. In this paper we describe
  performance measurements of Graphic Processing Units (GPUs) and Intel Many
  Integrated Core architecture (MIC) when applied to a typical HEP online task: 
  the selection of events based on the trajectories of charged particles.
  We use as benchmark a scaled-up version of the algorithm used at CDF 
  experiment at Tevatron for online track reconstruction - the SVT algorithm - 
  as a realistic  test-case for low-latency trigger systems using new computing
  architectures for LHC experiment. We examine the
  complexity/performance trade-off in porting existing serial
  algorithms to many-core devices. Measurements of both data processing 
  and data transfer latency are shown, considering different I/O strategies 
  to/from the parallel devices. \end{abstract}

@@

\subsection{Data processing in \textsc{opencl}}
RESULTS FROM THE FERMILAB SETUP.

>>>>

% disable the running header on all pages following the title page
\pagestyle{empty}
%
\begin{document}
% paper title
\title{Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics
Experiments}
% Author list
\author{A.~Gianelle, 
  S.~Amerio, 
  D.~Bastieri, 
  M.~Corvo, 
  W.~Ketchum,
  T.~Liu, 
  A.~Lonardo, 
  D.~Lucchesi,
  S.~Poprocki, 
  R.~Rivera, 
  L.~Tosoratto,
  P.~Vicini
  and 
  P.~Wittich
\thanks{Manuscript received November 15, 2013.
This work was supported by the  US National Science Foundation, the US 
Department of Energy Office of Science and the Italian
Istituto Nazionale di Fisica Nucleare. This work was partially supported by the 
EU Framework Programme 7 project EURETILE under grant number 247846.}% <-this % stops a space
\thanks{A.~Gianelle, S.~Amerio, D.~Bastieri, M.~Corvo, and D.~Lucchesi are with INFN Padova.}% 
\thanks{W.~Ketchum is with Los Alamos National Laboratory.}%
\thanks{T.~Liu and R.~Rivera are with Fermi National Accelerator Laboratory.}%
\thanks{A.~Lonardo, L.~Tosoratto and P.~Vicini are with INFN Roma.}% 
\thanks{S.~Poprocki and P.~Wittich are with Cornell University.}
}
% make the title area
\maketitle
% after \maketitle, add the following command to remove the header on the title page
\thispagestyle{empty}
% abstract usually in passive voice
\begin{abstract}
  Interest in parallel architectures applied to real time selections is 
  growing in High Energy Physics (HEP) experiments. In this paper we describe
  performance measurements of Graphic Processing Units (GPUs) and Intel Many
  Integrated Core architecture (MIC) when applied to a typical HEP online task: 
  the selection of events based on the trajectories of charged particles.
  We use as benchmark a scaled-up version of the algorithm used at CDF 
  experiment at Tevatron for online track reconstruction - the SVT algorithm - 
  as a realistic  test-case for low-latency trigger systems using new computing
  architectures for LHC experiment. We examine the
  complexity/performance trade-off in porting existing serial
  algorithms to many-core devices. Measurements of both data processing 
  and data transfer latency are shown, considering different I/O strategies 
  to/from the parallel devices. \end{abstract}

@@

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an implementation  
using \textsc{opencl}. Due to limitations on the available space for storing data on the GPU, we only test the 
algorithm on up to 1000 events (\textit{i.e.} about 2 million combinations to fit). In Fig.~\ref{fig:opencl_timing} 
we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (8-core), and running the \textsc{opencl} algorithm
on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
a factor of 5---over running the algorithm serially on the CPU. Running the same \textsc{opencl} algorithm on the GPU provides an 
even greater speedup due to the increased number of cores and parallel threads that can be run, though there is additional latency to copy data 
into and out of the GPU that makes running on the GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/Timing_OpenCL}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}
